---
title: "Unterstützer des Klimaschutz-Volksbegehrens"
subtitle: "Auswertungen"
author: "Stefanie Braunisch"
date: "06-2020"
output:
  html_document:
    code_folding: show
    toc: yes
    toc_float: 
      collapsed: false
      smooth_scroll: false
---


```{r, echo=FALSE}
# CONFIG
user_name <- "qvvdata" # your Git username (only needed if
# you want to deploy to GH pages)
project_name <- "klimavolksbegehren" # adapt!
package_date <- "2020-06-29" # date of the CRAN snapshot that
# the checkpoint package uses
```

## Notes

This report was generated on `r Sys.time()`.

...

### R-Script & data

The preprocessing and analysis of the data was conducted in the [R project for statistical computing](https://www.r-project.org/). The RMarkdown script used to generate this document and all the resulting data can be downloaded [under this link](http://`r user_name`.github.io/`r project_name`/rscript.zip). Through executing `main.Rmd`, the herein described process can be reproduced and this document can be generated. In the course of this, data from the folder `ìnput` will be processed and results will be written to `output`. 

### GitHub

The code for the herein described process can also be freely downloaded from [https://github.com/`r user_name`/`r project_name`](https://github.com/`r user_name`/`r project_name`). 


### License

...

### `xyz.csv` 

...

## Preparations

```{r, echo=FALSE}
detach_all_packages <- function() {
  basic_packages_blank <-  c("stats",
                             "graphics",
                             "grDevices",
                             "utils",
                             "datasets",
                             "methods",
                             "base")
  basic_packages <- paste("package:", basic_packages_blank, sep = "")

  package_list <- search()[
    ifelse(unlist(gregexpr("package:", search())) == 1, TRUE, FALSE)]

  package_list <- setdiff(package_list, basic_packages)

  if (length(package_list) > 0)  for (package in package_list) {
    detach(package, character.only = TRUE, unload = TRUE)
    print(paste("package ", package, " detached", sep = ""))
  }
}

detach_all_packages()

# this allows multiple persons to use the same RMarkdown
# without adjusting the working directory by themselves all the time
source("scripts/csf.R")
path_to_wd <- csf() # if this - for some reason - does not work, 
# replace with a hardcoded path, like so: "~/projects/rddj-template/analysis/"
if ( is.null(path_to_wd) | !dir.exists(path_to_wd)) {
  print("WARNING: No working directory specified for current user")
} else {
  setwd(path_to_wd)
}
```


### Define packages

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# from https://mran.revolutionanalytics.com/web/packages/checkpoint/vignettes/using-checkpoint-with-knitr.html
# if you don't need a package, remove it from here (commenting is probably not sufficient)
# tidyverse: see https://blog.rstudio.org/2016/09/15/tidyverse-1-0-0/
cat("
library(tidyverse) # ggplot2, dplyr, tidyr, readr, purrr, tibble
library(magrittr) # pipes
library(stringr) # string manipulation
library(readxl) # excel
library(scales) # scales for ggplot2
library(jsonlite) # json
library(forcats) # easier factor handling,
library(lintr) # code linting
library(openxlsx) #Excel
library(googlesheets) # googlesheets (replace with googlesheets4 asap)",
library(DatawRappr),
file = "manifest.R")
```

### Install packages

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# if checkpoint is not yet installed, install it (for people using this
# system for the first time)
if (!require(checkpoint)) {
  if (!require(devtools)) {
    install.packages("devtools", repos = "http://cran.us.r-project.org")
    require(devtools)
  }
  devtools::install_github("checkpoint",
                           username = "RevolutionAnalytics",
                           ref = "v0.3.2", # could be adapted later,
                           # as of now (beginning of July 2017
                           # this is the current release on CRAN)
                           repos = "http://cran.us.r-project.org")
  require(checkpoint)
}
# nolint start
if (!dir.exists("~/.checkpoint")) {
  dir.create("~/.checkpoint")
}
# nolint end
# install packages for the specified CRAN snapshot date
# checkpoint(snapshotDate = package_date,
#           project = path_to_wd,
#           verbose = T,
#           scanForPackages = T,
#           use.knitr = F)
rm(package_date)
```


### Load packages

```{r, echo=TRUE, message=FALSE, warning=FALSE}
#source("manifest.R")
unlink("manifest.R")
sessionInfo()
```

### Load additional scripts

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# if you want to outsource logic to other script files, see README for 
# further information
knitr::read_chunk("scripts/theme_addendum.R")
#source("./scripts/theme_addendum.R")
#source("./scripts/BorderMan.R")
source("./scripts/preprocessing_klimavolksbegehren.R")
cat(getwd())
#my_function(5)


#gemeindezusammenlegungen <- read.xlsx('input/2018-gemeindezusammenlegungen.xlsx', 1)
```



## Klimavolksbegehren


```{r, echo=TRUE, message=FALSE, warning=FALSE}
options(scipen=999)

#Analyse
analysis_rvb <- klimavolksbegehren %>%
            subset(klasse == "gem")

analysis_rvb <- analysis_rvb %>% left_join(contextdata, by=c("GKZ", "GKZ")) %>%
    mutate(
      pct = Summe/Wahlberechtigte*100, 
      pct_u = Unterstützungen/Wahlberechtigte*100,
      pct_e = Eintragungen/Wahlberechtigte*100,
      quantile = ntile(pct, 100),
      pct_diff = pct-pct_u,
      pct_plus = Unterstützungen/Summe*100)

analysis_rvb_r <- analysis_rvb %>%
  group_by(rurb) %>%
  summarise(ssum = sum(Summe), 
           vwb = sum(Wahlberechtigte),
          usum=sum(Unterstützungen)) %>%
  mutate(`Anteil der Unterstützer` = (usum/vwb)*100,
        `Anteil Gesamt` = (ssum/vwb)*100,
       Steigerung=`Anteil Gesamt`-`Anteil der Unterstützer`)

analysis_rvb_t <- analysis_rvb %>%
  group_by() %>%
  summarise(ssum = sum(Summe), 
            vwb = sum(Wahlberechtigte),
            usum=sum(Unterstützungen)) %>%
  mutate(`Anteil der Unterstützer` = (usum/vwb)*100,
         `Anteil Gesamt` = (ssum/vwb)*100,
         Steigerung=`Anteil Gesamt`-`Anteil der Unterstützer`)%>%
  mutate(typ = "Österreich gesamt") %>% 
  rename(rurb=typ)

analysis_rvb_r$rurb[analysis_rvb_r$rurb=="1"] <-"Urbane Zentren"
analysis_rvb_r$rurb[analysis_rvb_r$rurb=="2"] <-"Regionale Zentren"
analysis_rvb_r$rurb[analysis_rvb_r$rurb=="3"] <-"Ländliche Vororte"
analysis_rvb_r$rurb[analysis_rvb_r$rurb=="4"] <-"Ländlicher Raum"

analysis_rvb_r <- plyr::rbind.fill(analysis_rvb_r,analysis_rvb_t) %>% 
  select(rurb,ssum, vwb, usum, `Anteil der Unterstützer` , `Anteil Gesamt`, Steigerung)
  

write.csv2(analysis_rvb_r, "output/datawrapper/analysis_r.csv")

analysis_rvb_gem <- analysis_rvb %>%
  rename(gültig=gültige) %>% 
  mutate(tkgr = (GRÜNE+ÖVP)/gültig*100,
         gr= GRÜNE/gültig*100,
         tk=ÖVP/gültig*100) %>%
  select(GKZ, pct, pct_diff, Summe, Wahlberechtigte, pct, Name, tkgr,gr, tk, rurb)%>%
  rename(Anteil = pct, 
         Unterschriften = Summe,
         `türkis/grün-Anteil` = tkgr,
         `türkis Anteil` = tk,
         `grün Anteil`= gr)%>%
  mutate(Anteil = round(Anteil, 2))


#Hier werden nun klassen für analysis_gem zugewiesen

einheiten <- seq(0.1,1,0.1)

teiler_rvb <- quantile(analysis_rvb_gem$Anteil,probs=einheiten)

analysis_rvb_gem$cut <- cut(analysis_rvb_gem$Anteil,
                        breaks=c(0,teiler_rvb),
                        labels=paste(seq(10,100,10),"%",sep=""),
                        include.lowest=TRUE)

analysis_rvb_gem$klasse <- analysis_rvb_gem$cut

analysis_rvb_gem <- analysis_rvb_gem %>%
  mutate(klasse = gsub('%', '', klasse))%>%
  numerize(vars = c("Name", "cut", "urbanländlich"))%>%
  mutate(klasse =  if_else(klasse <=30, "unterdurchschnittlich",
                                   if_else(klasse >= 40 & klasse <=60, "durchschnittlich",
                                           if_else(klasse >= 60, "überdurchschnittlich","n.v.")))) 



#Bundesländer berechnen
bl_rvb  <- klimavolksbegehren %>%
            subset(klasse == "bl" | GKZ == "90001")%>%
  mutate(pct= Summe/Wahlberechtigte*100)

#Wiener Bezirke
wbez_rvb <- analysis_rvb_gem %>%
            subset(GKZ>90001)%>%
  mutate(pct= Unterschriften/Wahlberechtigte*100)%>%
  select(c(Name, pct, `grün Anteil`))

##Bezirke berechnen - eher unspektakulär
bz_rvb <- klimavolksbegehren %>% 
  subset(klasse=="bez") %>% 
  mutate(pct=Summe/Wahlberechtigte*100)

```
```{r}
DW_API_KEY <- "68cfebfe1fa4a7a517b9d8fd806392c4fceac44c7a35cc72c1298b708e54d465"
print('bkYTF')
gruen_Anteil <- analysis_rvb_gem %>% 
  select(Name, Anteil,`grün Anteil`) %>%
  rename(`Anteil Unterstützer Volksbegehren`=Anteil, `Anteil Grünwähler`=`grün Anteil`) %>% 
  dw_data_to_chart(chart_id = 'bkYTF', api_key = DW_API_KEY)
  

print('HENaC')
tuerkis_Anteil <- analysis_rvb_gem %>% 
  select(Name, Anteil,`türkis Anteil`) %>%
  rename(`Anteil Unterstützer Volksbegehren`=Anteil, `Anteil ÖVP-Wähler`=`türkis Anteil`) %>% 
  dw_data_to_chart(chart_id = 'HENaC', api_key = DW_API_KEY)

##Karte - Anteil von Wählern
print('7Bi6D')
Karte <- analysis_rvb_gem %>% 
  select(Name,Anteil, GKZ) %>% 
  dw_data_to_chart(chart_id='7Bi6D', api_key=DW_API_KEY)

###urban_rural
print('q903u')
Urban <- analysis_rvb_r [-1,] %>% 
  select(rurb, `Anteil der Unterstützer`,`Anteil Gesamt`, `Steigerung`) %>% 
  dw_data_to_chart(chart_id ='q903u', api_key = DW_API_KEY )
write_csv(analysis_rvb_r [-4, ], "output/datawrapper/urban.csv")

##Auswertung Wien
print('EH1y7')
wbez_rvb %>% 
  rename(`Anteil Unterstützer Volksbegehren`=pct, `Anteil Grünwähler`=`grün Anteil`) %>% 
  dw_data_to_chart(chart_id = 'EH1y7', api_key = DW_API_KEY)

```


```{r}
corrdf_rvb <- analysis_rvb %>%
  rename(gültig=gültige) %>% 
  mutate(fpoepct = FPÖ/gültig*100, 
         spoepct = SPÖ/gültig*100, 
         oevppct = ÖVP/gültig*100, 
         gruenepct = GRÜNE/gültig*100, 
         neospct= NEOS/gültig*100,
         tkgr = (ÖVP+GRÜNE)/gültig*100, 
         spfpneo = (SPÖ+FPÖ+NEOS)/gültig*100, 
         wb = Summe/gültig*100,
         GKZ=as.numeric(GKZ)) %>%
select(GKZ, pct, fpoepct, spoepct, oevppct, gruenepct, tkgr, spfpneo, wb) %>%
  na.omit()

corrdf_rvb <- cor(corrdf_rvb, use="pairwise.complete.obs")
round(corrdf_rvb, digits=2)

corrdf_rvb <- as.data.frame(corrdf_rvb)
corrdf2018_rvb <- corrdf_rvb %>%
            subset(select=c(oevppct, spoepct, gruenepct, fpoepct))
```



```{r}
bl <- klimavolksbegehren %>%
  arrange(desc(Wahlberechtigte)) %>%
  top_n(10, Wahlberechtigte) %>%
  mutate(pct = Unterstützungen/Wahlberechtigte*100)%>%
  arrange(desc(pct))
write.csv(bl, "output/datawrapper/bl.csv")


#Wiener Bezirke
wbez_rvb <- klimavolksbegehren %>%
            subset(GKZ>90001)%>%
  mutate(pct= Summe/Wahlberechtigte*100)%>%
  select(c(Name, pct))
write.csv(wbez_rvb, "output/datawrapper/wbez.csv")

#Visualisierungen

  # labs(title = "Schlagzeile", 
  #      x = "Vertikale Achse", 
  #      y = "horizontale Achse",
  #      #subtitle ="Funktioniert aus unerklärlichen Gründen noch nicht", 
  #      caption = "Quelle")+
  # theme_addendum()

#Exportieren für Illustrator funktioniert nicht über ggsave (Schriften), sondern besser über Export des Plot-Viewers
```



## Linting

The code in this RMarkdown is listed with the [lintr package](https://github.com/jimhester/lintr), which is based on the  [tidyverse style guide](http://style.tidyverse.org/). 

```{r echo=TRUE, message=FALSE, warning=FALSE}
lintr::lint("main.Rmd")
# if you have additional scripts and want them to be linted too, add them here
#lintr::lint("scripts/my_script.R")
```


