---
title: "Unterstützer des Nichtraucher-Volksbegehrens"
subtitle: "Auswertungen"
author: "Gerald Gartner"
date: "03-2018"
output:
  html_document:
    code_folding: show
    toc: yes
    toc_float: 
      collapsed: false
      smooth_scroll: false
---


```{r, echo=FALSE}
# CONFIG
user_name <- "qvvdata" # your Git username (only needed if
# you want to deploy to GH pages)
project_name <- "nichtrauchervolksbegehren" # adapt!
package_date <- "2018-03-03" # date of the CRAN snapshot that
# the checkpoint package uses
```

## Notes

This report was generated on `r Sys.time()`.

...

### R-Script & data

The preprocessing and analysis of the data was conducted in the [R project for statistical computing](https://www.r-project.org/). The RMarkdown script used to generate this document and all the resulting data can be downloaded [under this link](http://`r user_name`.github.io/`r project_name`/rscript.zip). Through executing `main.Rmd`, the herein described process can be reproduced and this document can be generated. In the course of this, data from the folder `ìnput` will be processed and results will be written to `output`. 

### GitHub

The code for the herein described process can also be freely downloaded from [https://github.com/`r user_name`/`r project_name`](https://github.com/`r user_name`/`r project_name`). 


### License

...

### `xyz.csv` 

...

## Preparations

```{r, echo=FALSE}
detach_all_packages <- function() {
  basic_packages_blank <-  c("stats",
                             "graphics",
                             "grDevices",
                             "utils",
                             "datasets",
                             "methods",
                             "base")
  basic_packages <- paste("package:", basic_packages_blank, sep = "")

  package_list <- search()[
    ifelse(unlist(gregexpr("package:", search())) == 1, TRUE, FALSE)]

  package_list <- setdiff(package_list, basic_packages)

  if (length(package_list) > 0)  for (package in package_list) {
    detach(package, character.only = TRUE, unload = TRUE)
    print(paste("package ", package, " detached", sep = ""))
  }
}

detach_all_packages()

# this allows multiple persons to use the same RMarkdown
# without adjusting the working directory by themselves all the time
source("scripts/csf.R")
path_to_wd <- csf() # if this - for some reason - does not work, 
# replace with a hardcoded path, like so: "~/projects/rddj-template/analysis/"
if ( is.null(path_to_wd) | !dir.exists(path_to_wd)) {
  print("WARNING: No working directory specified for current user")
} else {
  setwd(path_to_wd)
}
```


### Define packages

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# from https://mran.revolutionanalytics.com/web/packages/checkpoint/vignettes/using-checkpoint-with-knitr.html
# if you don't need a package, remove it from here (commenting is probably not sufficient)
# tidyverse: see https://blog.rstudio.org/2016/09/15/tidyverse-1-0-0/
cat("
library(tidyverse) # ggplot2, dplyr, tidyr, readr, purrr, tibble
library(magrittr) # pipes
library(stringr) # string manipulation
library(readxl) # excel
library(scales) # scales for ggplot2
library(jsonlite) # json
library(forcats) # easier factor handling,
library(lintr) # code linting
library(xlsx) #Excel
library(googlesheets) # googlesheets (replace with googlesheets4 asap)",
file = "manifest.R")
```

### Install packages

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# if checkpoint is not yet installed, install it (for people using this
# system for the first time)
if (!require(checkpoint)) {
  if (!require(devtools)) {
    install.packages("devtools", repos = "http://cran.us.r-project.org")
    require(devtools)
  }
  devtools::install_github("checkpoint",
                           username = "RevolutionAnalytics",
                           ref = "v0.3.2", # could be adapted later,
                           # as of now (beginning of July 2017
                           # this is the current release on CRAN)
                           repos = "http://cran.us.r-project.org")
  require(checkpoint)
}
# nolint start
if (!dir.exists("~/.checkpoint")) {
  dir.create("~/.checkpoint")
}
# nolint end
# install packages for the specified CRAN snapshot date
# checkpoint(snapshotDate = package_date,
#           project = path_to_wd,
#           verbose = T,
#           scanForPackages = T,
#           use.knitr = F)
rm(package_date)
```


### Load packages

```{r, echo=TRUE, message=FALSE, warning=FALSE}
source("manifest.R")
unlink("manifest.R")
sessionInfo()
```

### Load additional scripts

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# if you want to outsource logic to other script files, see README for 
# further information
knitr::read_chunk("scripts/theme_addendum.R")
source("./scripts/theme_addendum.R")
source("./scripts/BorderMan.R")
source("./scripts/01-preprocessing.R")
cat(getwd())
#my_function(5)


gemeindezusammenlegungen <- read.xlsx('input/2018-gemeindezusammenlegungen.xlsx', 1)
```



## Rauchervolksbegehren


```{r, echo=TRUE, message=FALSE, warning=FALSE}
options(scipen=999)

#Analyse
analysis_rvb <- rauchervolksbegehren %>%
            subset(klasse == "gem")

analysis_rvb <- analysis_rvb %>% left_join(contextdata, by=c("gkz", "gkz")) %>%
    mutate(
      pct = Summe/Wahlberechtigte*100, 
      pct_u = Unterstützungen/Wahlberechtigte*100,
      pct_e = Eintragungen/Wahlberechtigte*100,
      quantile = ntile(pct, 100),
      pct_diff = pct-pct_u,
      pct_plus = Unterstützungen/Summe*100)

analysis_rvb_r <- analysis_rvb %>%
  group_by(typ) %>%
  summarise(ssum = sum(Summe), 
            vwb = sum(Wahlberechtigte),
            usum=sum(Unterstützungen)) %>%
  mutate(`Anteil der Unterstützer` = (usum/vwb)*100,
         `Anteil Gesamt` = (ssum/vwb)*100,
         Steigerung=`Anteil Gesamt`-`Anteil der Unterstützer`)

analysis_rvb_t <- analysis_rvb %>%
  group_by() %>%
  summarise(ssum = sum(Summe), 
            vwb = sum(Wahlberechtigte),
            usum=sum(Unterstützungen)) %>%
  mutate(`Anteil der Unterstützer` = (usum/vwb)*100,
         `Anteil Gesamt` = (ssum/vwb)*100,
         Steigerung=`Anteil Gesamt`-`Anteil der Unterstützer`)%>%
  mutate(typ = "Österreich gesamt")

analysis_rvb_r$typ[analysis_rvb_r$typ=="1"] <-"Urbane Zentren"
analysis_rvb_r$typ[analysis_rvb_r$typ=="2"] <-"Regionale Zentren"
analysis_rvb_r$typ[analysis_rvb_r$typ=="3"] <-"Ländliche Vororte"
analysis_rvb_r$typ[analysis_rvb_r$typ=="4"] <-"Ländlicher Raum"

analysis_rvb_r <- plyr::rbind.fill(analysis_rvb_r,analysis_rvb_t)


write.csv2(analysis_rvb_r, "output/datawrapper/analysis_r.csv")

analysis_rvb_gem <- analysis_rvb %>%
  mutate(tkbl = (fpoe+oevp)/gültig*100) %>%
  select(gkz, pct, pct_diff, Summe, Wahlberechtigte, pct, Name, tkbl, `urbanländlich`)%>%
  rename(
         Anteil = pct, 
         Unterschriften = Summe,
         `türkis/blau-Anteil` = tkbl) %>%
  mutate(Anteil = round(Anteil, 2))


#Hier werden nun klassen für analysis_gem zugewiesen

einheiten <- seq(0.1,1,0.1)

teiler_rvb <- quantile(analysis_rvb_gem$Anteil,probs=einheiten)

analysis_rvb_gem$cut <- cut(analysis_rvb_gem$Anteil,
                        breaks=c(0,teiler_rvb),
                        labels=paste(seq(10,100,10),"%",sep=""),
                        include.lowest=TRUE)

analysis_rvb_gem$klasse <- analysis_rvb_gem$cut

analysis_rvb_gem <- analysis_rvb_gem %>%
  mutate(klasse = gsub('%', '', klasse))%>%
  numerize(vars = c("Name", "cut", "urbanländlich"))%>%
  mutate(klasse =  if_else(klasse <=30, "unterdurchschnittlich",
                                   if_else(klasse >= 40 & klasse <=60, "durchschnittlich",
                                           if_else(klasse >= 60, "überdurchschnittlich","n.v.")))) %>%
  rename(gkz_neu = gkz)
  

write.csv(analysis_rvb_gem %>%
  select(gkz_neu, Anteil, Unterschriften, pct_diff, Wahlberechtigte,Name) %>%
    left_join(gemeindezusammenlegungen) %>% mutate(gkz_neu = case_when(is.na(gkz_alt) ~ gkz_neu, TRUE ~ gkz_alt)) %>% select(-name_alt,-gkz_alt), "interaktiv/karte/dist/analysis_gem.csv",fileEncoding='UTF-8',row.names = F)


write.csv2(analysis_rvb_gem %>% select(gkz_neu, Anteil, `türkis/blau-Anteil`, `urbanländlich`, Unterschriften) %>% rename(tkbl=`türkis/blau-Anteil`, gkz=gkz_neu, pct=Anteil, Summe=Unterschriften), "output/datawrapper/analysis_rvb_gem.csv")



#Bundesländer berechnen
bl_rvb  <- rauchervolksbegehren %>%
            subset(klasse == "bl" | gkz == "90001")%>%
  mutate(pct= Summe/Wahlberechtigte*100)

#Wiener Bezirke
wbez_rvb <- rauchervolksbegehren %>%
            subset(gkz>90001)%>%
  mutate(pct= Summe/Wahlberechtigte*100)%>%
  select(c(Name, pct))
write.csv(wbez_rvb, "output/datawrapper/wbez.csv")



```


```{r}
corrdf_rvb <- analysis_rvb %>%
  mutate(fpoepct = fpoe/gültig*100, 
         spoepct = spoe/gültig*100, 
         oevppct = oevp/gültig*100, 
         gruenepct = gruene/gültig*100, 
         tkbl = (fpoe+oevp)/gültig*100, 
         spgrneo = (spoe+gruene+neos)/gültig*100, 
         wb = abgegeben/gültig*100) %>%
select(gkz, pct, fpoepct, spoepct, oevppct, gruenepct, tkbl, spgrneo, wb) %>%
  na.omit()

corrdf_rvb <- cor(corrdf_rvb, use="pairwise.complete.obs")
round(corrdf_rvb, digits=2)

corrdf_rvb <- as.data.frame(corrdf_rvb)
corrdf2018_rvb <- corrdf_rvb %>%
            subset(select=c(oevppct, spoepct, gruenepct, fpoepct))
```


#Frauenvolksbegehren analysieren

```{r}
#Analyse
analysis_fvb <- fvb_8_3 %>%
            subset(klasse == "gem")

analysis_fvb <- analysis_fvb %>% left_join(contextdata, by=c("gkz", "gkz")) %>%
    mutate(
      pct = Summe/Wahlberechtigte*100, 
      pct_u = Eintragungen/Wahlberechtigte*100,
      Steigerung = pct-pct_u,
      quantile = ntile(pct, 100))

analysis_fvb_r <- analysis_fvb %>%
  group_by(typ) %>%
  summarise(usum = sum(`Unterstützungen`), 
            ssum = sum(Summe),
            vwb = sum(Wahlberechtigte)) %>%
  mutate(`Anteil der Unterstützer` = (usum/vwb)*100,
         `Anteil gesamt` = (ssum/vwb)*100,
        Steigerung = `Anteil gesamt`-`Anteil der Unterstützer`)

analysis_fvb_t <- analysis_fvb %>%
  group_by() %>%
  summarise(usum = sum(`Unterstützungen`), 
            ssum = sum(Summe),
            vwb = sum(Wahlberechtigte)) %>%
  mutate(`Anteil der Unterstützer` = (usum/vwb)*100,
         `Anteil gesamt` = (ssum/vwb)*100,
        Steigerung = `Anteil gesamt`-`Anteil der Unterstützer`)%>%
  mutate(typ = "Österreich gesamt")

analysis_fvb_r$typ[analysis_fvb_r$typ=="1"] <-"Urbane Zentren"
analysis_fvb_r$typ[analysis_fvb_r$typ=="2"] <-"Regionale Zentren"
analysis_fvb_r$typ[analysis_fvb_r$typ=="3"] <-"Ländliche Vororte"
analysis_fvb_r$typ[analysis_fvb_r$typ=="4"] <-"Ländlicher Raum"

analysis_fvb_r <- plyr::rbind.fill(analysis_fvb_r,analysis_fvb_t)


write.csv2(analysis_fvb_r, "output/datawrapper/analysis_fvb_r.csv")

analysis_fvb_gem <- analysis_fvb %>%
  mutate(tkbl = (fpoe+oevp)/gültig*100) %>%
  select(gkz, Summe, Unterstützungen, Wahlberechtigte, Steigerung, pct, Name, tkbl)%>%
  rename( 
         Anteil = pct, 
         `türkis/blau-Anteil` = tkbl) %>%
  mutate(Anteil = round(Anteil, 2))%>%
  select(gkz, Anteil, Summe, Unterstützungen, Wahlberechtigte, Steigerung, Name)
write.csv2(analysis_fvb_gem, "output/datawrapper/analysis_fvb_gem.csv")

#Bundesländer berechnen
bl_fvb  <- fvb_8_3 %>%
            subset(klasse == "bl" | gkz == "90001")%>%
  mutate(pct=Summe/Wahlberechtigte*100)

#Wiener Bundesländer
wbez_fvb <- fvb_8_3 %>%
            subset(gkz>90001)%>%
  mutate(pct=Summe/Wahlberechtigte*100)%>%
  select(c(Name, pct))
write.csv2(wbez_fvb, "output/datawrapper/wbez_fvb.csv")

#Bezirksanalyse mit 1994
fvbbez <- fvb_8_3 %>%
            subset(klasse =="bez" | gkz=="90001")%>%
          mutate(bkz = substr(gkz, 1, 3))%>%
  numerize(vars = c("Name", "klasse")) 

fvbbez_altneu <- fvb97 %>% left_join(fvbbez, by=c("bkz", "bkz")) %>%
  na.omit() %>%
    mutate(
      pct97 = frauengueltig/wb*100, 
      pct18 = Summe/Wahlberechtigte*100)%>%
      select(Name, pct97, pct18, Wahlberechtigte,lh)

write.csv2(fvbbez_altneu, "output/datawrapper/fvbbez_altneu.csv")

write.csv2(analysis_fvb_gem, "output/datawrapper/analysis_fvb_gem_alle.csv")

fvb97
```



```{r}
bl <- rauchervolksbegehren %>%
  arrange(desc(Wahlberechtigte)) %>%
  top_n(10, Wahlberechtigte) %>%
  mutate(pct = Unterstützungen/Wahlberechtigte*100)%>%
  arrange(desc(pct))
write.csv(bl, "output/datawrapper/bl.csv")

#Visualisierungen

  # labs(title = "Schlagzeile", 
  #      x = "Vertikale Achse", 
  #      y = "horizontale Achse",
  #      #subtitle ="Funktioniert aus unerklärlichen Gründen noch nicht", 
  #      caption = "Quelle")+
  # theme_addendum()

#Exportieren für Illustrator funktioniert nicht über ggsave (Schriften), sondern besser über Export des Plot-Viewers
```

#ORF VOLKSBEGEHREN ANALYSEN
```{r}
#Analyse
analysis_orf <- orf %>%
            subset(klasse == "gem")

analysis_orf <- analysis_orf %>% left_join(contextdata, by=c("gkz", "gkz")) %>%
    mutate(
      pct = Summe/Wahlberechtigte*100,
      pctu = Unterstützungen/Wahlberechtigte*100, 
      pcte = Eintragungen/Wahlberechtigte*100,
      quantile = ntile(pct, 100))

analysis_orf_r <- analysis_orf %>%
  group_by(typ) %>%
  summarise(usum = sum(Unterstützungen),
            esum = sum(`Eintragungen`),
            sumsum = sum(Summe), 
            vwb = sum(Wahlberechtigte)) %>%
  mutate(`Anteil der Unterstützer` = (usum/vwb)*100, 
         `Anteil der Eintragungen` = (esum/vwb)*100,
         `Anteil gesamt` = (sumsum/vwb)*100, 
          Steigerung = `Anteil gesamt`-`Anteil der Unterstützer`)

analysis_orf_t <- analysis_orf %>%
  group_by() %>%
  summarise(usum = sum(Unterstützungen), 
            esum = sum(`Eintragungen`),
            sumsum = sum(Summe),
            vwb = sum(Wahlberechtigte)) %>%
  mutate(
        `Anteil der Unterstützer` = (usum/vwb)*100, 
         `Anteil der Eintragungen` = (esum/vwb)*100,
         `Anteil gesamt` = (sumsum/vwb)*100, 
          Steigerung = `Anteil gesamt`-`Anteil der Unterstützer`)%>%
  mutate(typ = "Österreich gesamt")

analysis_orf_r$typ[analysis_orf_r$typ=="1"] <-"Urbane Zentren"
analysis_orf_r$typ[analysis_orf_r$typ=="2"] <-"Regionale Zentren"
analysis_fvb_r$typ[analysis_orf_r$typ=="3"] <-"Ländliche Vororte"
analysis_orf_r$typ[analysis_orf_r$typ=="4"] <-"Ländlicher Raum"

analysis_orf_r <- plyr::rbind.fill(analysis_orf_r,analysis_orf_t)


write.csv2(analysis_orf_r, "output/datawrapper/analysis_orf_r.csv")

analysis_orf_gem <- analysis_orf %>%
  mutate(tkbl = (fpoe+oevp)/gültig*100, 
         Steigerung = pct-pctu) %>%
  select(gkz, Eintragungen, Summe, Unterstützungen, Wahlberechtigte, pct, pctu, pcte, Name, tkbl, Steigerung, pctu)%>%
  rename(
         `Anteil gesamt`= pct,
         `Anteil Unterstützungsphase` = pctu,
         Unterstützer = Summe, 
         `türkis/blau-Anteil` = tkbl
         ) %>%
  mutate(`Anteil gesamt` = round(`Anteil gesamt`, 2), 
         Steigerung = round(Steigerung, 2), 
         `Anteil Unterstützungsphase` = round(`Anteil Unterstützungsphase`,2))
write.csv2(analysis_orf_gem %>% select(gkz, `Anteil gesamt`, `Anteil Unterstützungsphase`, Unterstützer, Wahlberechtigte,Name, `türkis/blau-Anteil`, Steigerung)
, "output/datawrapper/analysis_orf_gem.csv")

#Bundesländer berechnen
bl_orf  <- orf %>%
            subset(klasse == "bl" | gkz == "90001")%>%
  mutate(pct= Summe/Wahlberechtigte*100)
write.csv2(bl_orf, "output/datawrapper/bl_orf.csv")

#Wiener bezirke
wbez_orf <- analysis_orf %>%
            subset(gkz>90001)%>%
  mutate(pct= Summe/Wahlberechtigte*100)%>%
  select(c(Name, pct))
write.csv2(wbez_fvb, "output/datawrapper/wbez_orf.csv")


#write.csv2(analysis_orf_gem, "output/datawrapper/analysis_orf_gem_alle.csv") 
analysis_orf_gem_wien <- analysis_orf_gem %>% filter(gkz>90000) %>% summarise(
  gkz=90000,
  Eintragungen=sum(Eintragungen),
  Unterstützer=sum(Unterstützer),
  Unterstützungen=sum(Unterstützungen),
  Wahlberechtigte=sum(Wahlberechtigte),
  `Anteil gesamt`=round(Unterstützer/Wahlberechtigte*100,2),
  `Anteil Unterstützungsphase`=round(Unterstützungen/Wahlberechtigte*100,2),
  Steigerung=`Anteil gesamt`-`Anteil Unterstützungsphase`,
  Name="Wien"
)
write.csv(analysis_orf_gem %>% bind_rows(analysis_orf_gem_wien) %>% select(-pcte,-`türkis/blau-Anteil`) %>% mutate(`Anteil gesamt` = gsub('\\.',',',as.character(`Anteil gesamt`)))  %>%
    left_join(gemeindezusammenlegungen, by=c(gkz='gkz_neu')) %>% mutate(gkz = case_when(is.na(gkz_alt) ~ gkz, TRUE ~ gkz_alt)) %>% select(-name_alt,-gkz_alt), "output/datawrapper/analysis_orf_gem_alle.csv")


#Korrelationen
corrdf <- analysis_orf %>%
  mutate(fpoepct = fpoe/gültig*100, 
         spoepct = spoe/gültig*100, 
         oevppct = oevp/gültig*100, 
         gruenepct = gruene/gültig*100, 
         tkbl = (fpoe+oevp)/gültig*100,
         nw = (wb_nrw-abgegeben)/wb_nrw*100,
         spgrneo = (spoe+gruene+neos)/gültig*100, 
         wb = abgegeben/gültig*100) %>%
select(gkz, pct, fpoepct, spoepct, oevppct, gruenepct, tkbl, spgrneo, wb, nw) %>%
  na.omit()



write.csv(analysis_orf_gem %>%
  select(gkz, `Anteil gesamt`, Unterstützer, Steigerung, Wahlberechtigte,Name) %>%
    rename(gkz_neu = gkz, Anteil=`Anteil gesamt`, Unterschriften=Unterstützer, pct_diff=Steigerung) %>%
    left_join(gemeindezusammenlegungen) %>% mutate(gkz_neu = case_when(is.na(gkz_alt) ~ gkz_neu, TRUE ~ gkz_alt)) %>% select(-name_alt,-gkz_alt), "interaktiv/karte/dist/analysis_gem_orf.csv",fileEncoding='UTF-8',row.names = F)

corrdf <- cor(corrdf, use="pairwise.complete.obs")
round(corrdf, digits=2)

corrdf <- as.data.frame(corrdf)
corrdf2018 <- corrdf %>%
            subset(select=c(oevppct, spoepct, gruenepct, fpoepct))
```


## Linting

The code in this RMarkdown is listed with the [lintr package](https://github.com/jimhester/lintr), which is based on the  [tidyverse style guide](http://style.tidyverse.org/). 

```{r echo=TRUE, message=FALSE, warning=FALSE}
lintr::lint("main.Rmd")
# if you have additional scripts and want them to be linted too, add them here
#lintr::lint("scripts/my_script.R")
```


